---
title: "Mexicana_Emma"
output: html_document
---

Main hypothesis: selection for genome size along an altitudinal gradient across the entire range.

```{r}
#install.packages("emma")
library ( emma )
#install.packages("rrBLUP")
library("rrBLUP")

<<<<<<< HEAD
source("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/SelectionTests/jri_emma_returnbeta.txt")

=======
>>>>>>> 9418a0eb3035598edffcb06d5c38aa669b8c3859
pheno <- read.csv("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Master_mexnucleo_pruned.csv") #phenotype data in matching order
geno <- read.csv("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/SNP_data/Mexicana_conversion/GBS_final.txt")
geno$X <- NULL

dt <- t(geno)

A <- A.mat(dt)

#create seperate matrices for each phenotype you want to model.
gs <- t ( as.matrix ( pheno$GS_bp, ncol = 1 ) )
gsgb <- gs/1000000000

knob180 <- t ( as.matrix (pheno$X180knobbp, ncol=1))
knob180gb <- knob180/1000000000

alt <- t ( as.matrix ( pheno$Altitude , ncol = 1 ) )
alt <- alt - mean ( alt ) + 0.5
#mean centering between 0 and 1

A <- as.matrix ( A )
#make sure your kinship matrix is a matrix.

<<<<<<< HEAD
out.gsgb <- emma.REML.t_beta ( gsgb , alt , K = A )
out.gsgb

gs <- t ( as.matrix ( pheno$X180knobbp, ncol = 1 ) )
gsgb <- gs/1000000000

out.gsgb <- emma.REML.t_beta ( gsgb , alt , K = A )
out.gsgb

gs <- t ( as.matrix ( pheno$TR1bp, ncol = 1 ) )
gsgb <- gs/1000000000

out.gsgb <- emma.REML.t_beta ( gsgb , alt , K = A )
out.gsgb

gs <- t ( as.matrix ( pheno$TotalTebp, ncol = 1 ) )
gsgb <- gs/1000000000

out.gsgb <- emma.REML.t_beta ( gsgb , alt , K = A )
out.gsgb


```

Include per individual inbreeding as a covariate, see if altitude term is then significant in regard to genome size.  
```{r}
hets<-read.csv("~/Desktop/teosinte_diversity/dothet.csv")
inbr <- t ( as.matrix ( hets$F , ncol = 1 ) )

fin.gsinbr <- emma.REML.t ( gsgb , alt , X0 =  cbind ( rep ( 1 , 93 ) , c ( inbr ) ) , K = A )
fin.gsinbr
```




Checking for knob patterns in low elevation mexicana
```{r}
par(mfrow=c(2,2))
plot(pheno$Altitude,pheno$X180knobbp,xlab="Altitude (m)",ylab="BP 180 knob")
plot(pheno$Altitude,pheno$X180knob.,xlab="Altitude (m)",ylab="% 180 knob")

plot(pheno$Altitude,pheno$TR1bp,xlab="Altitude (m)",ylab="BP TR1 knob")
plot(pheno$Altitude,pheno$TR1.,xlab="Altitude (m)",ylab="% TR1 knob")

=======
out.gsgb <- emma.REML.t ( gsgb , alt , K = A )
out.gsgb
>>>>>>> 9418a0eb3035598edffcb06d5c38aa669b8c3859
```

Alternate hypothesis.  Data cleaved at the 2200M Altitude to look for linear selection after that height threshold.

```{r}
library(emma)
library("rrBLUP")
setwd("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/SelectionTests/SelectionTestData/")

gbs<-read.csv("GBS_alt_threshold_final.txt")
gbs$X <- NULL
dt<-(t(gbs))
A<-A.mat(dt)
pheno<-read.csv("Pheno_alt_threshold_ordered.csv")

gs <- t ( as.matrix (pheno$GS_bp, ncol=1) )
gsgb <- gs/1000000000

alt <- t ( as.matrix (pheno$Altitude, ncol=1))
alt <- alt - mean(alt) + 0.5

<<<<<<< HEAD
knob180 <- t ( as.matrix (pheno$X180knobbp, ncol=1))
knob180gb <- knob180/1000000000

TR1 <- t ( as.matrix (pheno$TR1bp, ncol=1))
TR1gb <- TR1/1000000000

Te <- t ( as.matrix (pheno$TotalTebp, ncol=1))
Tegb <- Te/1000000000


out.gsgb <- emma.REML.t_beta( gsgb,alt,K=A)
out.gsgb

out.knob <- emma.REML.t_beta(knob180gb,alt,K=A)
out.knob

out.tr1 <- emma.REML.t_beta(TR1gb,alt,K=A)
out.tr1

out.te <- emma.REML.t_beta(Tegb,alt,K=A)
out.te

fin.gs <- emma.REML.t_beta ( gsgb , alt , X0 =  cbind ( rep ( 1 , 70 ) , c ( TR1gb ) ) , K = A )
fin.gs


fin.knob <- emma.REML.t_beta ( knob180gb , alt , X0 =  cbind ( rep ( 1 , 70 ) , c ( gsgb ) ) , K = A )
fin.knob

fin.tr1 <- emma.REML.t_beta ( TR1gb , alt , X0 =  cbind ( rep ( 1 , 70 ) , c ( gsgb ) ) , K = A )
fin.tr1

fin.te <- emma.REML.t_beta ( Tegb , alt , X0 =  cbind ( rep ( 1 , 70 ) , c ( gsgb ) ) , K = A )
fin.te


plot(pheno$Altitude,pheno$X1C_GS,xlab="Altitude (m)",ylab="2C Genome Size (pg)",pch=19, col="black")
abline(lm(pheno$X1C_GS~pheno$Altitude),lwd=2,lty=2)
?abline
plot(pheno$Altitude,pheno$X180knobbp,xlab="Altitude (m)",ylab="Knob abundance",pch=19, col="black")
plot(pheno$Altitude,pheno$TR1bp,xlab="Altitude (m)",ylab="TR1 abundance",pch=19, col="black")
plot(pheno$Altitude,pheno$TotalTebp,xlab="Altitude (m)",ylab="TE abundance",pch=19, col="black")

plot(pheno$Altitude,pheno$CentCbp,xlab="Altitude (m)",ylab="Centc abundance",pch=19, col="black")
=======
out.gsgb <- emma.REML.t( gsgb,alt,K=A)
out.gsgb
plot(pheno$Altitude,pheno$X1C_GS,xlab="Altitude (m)",ylab="2C Genome Size (pg)",pch=19, col="salmon")
abline(lm(pheno$X1C_GS~pheno$Altitude),lwd=2,lty=2)
?abline
>>>>>>> 9418a0eb3035598edffcb06d5c38aa669b8c3859
```

