---
title: "Final_Grote_growthchamber"
output: html_document
---

Identifying effect sizes in the two mediator model of genome size and leaf elongation.

Step1: Load in data and execute JAGS model to generate posterior distributions of components.
Step2: Use posterior distributions to generate figures to show confidence intervals on real data in the unit of measure.

```{r}
library(rjags)
library(stringr)
library(mcmcplots)
setwd("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Finalgrowthchamberanalysis/")
```

Load in data sets, build data frames necessary for JAGS input.

```{r}
#Master Data frame with all measures is called fgrote, for final grote dataframe.
fgrote<-read.csv("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Finalgrowthchamberanalysis/grote_Indexmoms_groundedLL_d16NA.csv")

#Make the necessary dataframe inputs into the model
stoma.positions <- str_detect(names(fgrote),"stom*")
stoma <- fgrote[,names(fgrote)[stoma.positions]]
stoma <- 0.0001 * stoma #convert mircons to centimeters
leaflength.positions <- str_detect(names(fgrote), "leaf3diff*")#change this to leaf3diff*
leaflength <- fgrote[,names(fgrote)[leaflength.positions]]
n.cells <- apply(!is.na(stoma), MAR=1, FUN="sum")

#data frames needed for figure generation
gensize.conforming <- matrix(fgrote$genomesize,nrow=nrow(fgrote),ncol=3,byrow=FALSE)
fgrote$leaf3diff10 <- fgrote$leaf3.1 - fgrote$leaf3.0
fgrote$leaf3diff21 <- fgrote$leaf3.2 - fgrote$leaf3.1
fgrote$leaf3diff32 <- fgrote$leaf3.3 - fgrote$leaf3.2
LEdiffs <- as.matrix(cbind(fgrote$leaf3diff10,fgrote$leaf3diff21,fgrote$leaf3diff32))
#emperical derivates, change per day, necessary for figures
```

Read in, burn in, and execute the model.  This segment will take time, depending on how many iterations you want to calculate and store.

```{r}
DAGLL.jags <- jags.model("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Finalgrowthchamberanalysis/DAG_twomediator_Final.txt", list(leaflength=log(leaflength),stoma=log(stoma),n.mothers=max(fgrote$momindex),n.indiv=dim(fgrote)[1],n.cells=n.cells,n.times=3,mother.index=fgrote$momindex,GS=log(fgrote$genomesize)),n.chains=2)

update(DAGLL.jags, n.iter=200000)

DAGLL.samples <- coda.samples(DAGLL.jags, c("beta.GS","tao.GS","GS.contrast","sd.motherStCS","sd.motherLL","sd.indivCS","sd.indivLL","sd.StCS","sd.LL","StCS.coeffs","tao.LL","LE.coeffs"),n.iter=500000, thin=500)

mcmcplot(DAGLL.samples)
```

Save chains, chunk not to be executed each time.

```{r}
chain1<- as.matrix(DAGLL.samples[[1]])
chain2<- as.matrix(DAGLL.samples[[2]])
write.csv(chain1,"DAGLLfinal.jags.le.6_003_4.8.chain1.csv")
write.csv(chain2,"DAGLLfinal.jags.le.6_003_4.8.chain2.csv")
```

Plot genome size by leaf elongation.

```{r}
#placeholder
fgrote<-read.csv("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Finalgrowthchamberanalysis/grote_Indexmoms_groundedLL_d16NA.csv")
tmp1<- read.csv("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Grote_stats/Gridcomparison/gamma.chain2.ind4.8mom0_indcellsize.csv")
tmp2<- read.csv("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Grote_stats/Gridcomparison/gamma.chain1.ind4.8mom0_indcellsize.csv")
tmp1 <- mcmc(tmp1,thin=500)
tmp2 <- mcmc(tmp2,thin=500)
#samples2 <- mcmc.list(chain1=tmp1,chain2=tmp2)
samples<-rbind(tmp1,tmp2)

colnames(samples)

#plot(samples.mat[,"tao.GS"],samples.mat[,"LE.coeffs.143."])


#Set up plotting boundaries of the graph
loggs.seq <- seq(from=min(log(fgrote$genomesize)),to=max(log(fgrote$genomesize)), by=0.001)

loggs.mat <- cbind(rep(1,length(loggs.seq)),rep(1,length(loggs.seq)), loggs.seq) #matrix with 3 columns, first 2 columns are =1 to pick up individual random effect for leaf length and leaf elongation, and the third is a sequence for the axis of log genome size

gs.params <- samples[,c("U.z.indivLL","tao.GS")]
gs.params <- cbind(log(samples$U.z.indivLE),gs.params)#collect the realizations (intercepts) that correspond to the effect being displayed
coef.means <- apply(gs.params, MAR=2, FUN="mean") #take the posterior mean for the central tendency curve

#gs.params needs to have columns for each of the realization terms, might include average(z.indivll+z.momll) or might have other terms.  need to add in a rep(1,length(loggs.seq) for each additional terms

predmean <- exp(loggs.mat %*% coef.means) #matrix math for the central tendency curve

#making of the credibility bands

allpreds <- as.matrix(gs.params) %*% t(loggs.mat) #evaluate the linear predictor for every realization
allpreds <- exp(allpreds) 
#dim(allpreds)
preds.upper <- apply(allpreds,MAR=2,FUN="quantile",probs=c(0.975)) #upper bounds of our predictor
preds.lower <- apply(allpreds,MAR=2,FUN="quantile",probs=c(0.025))

ymin <- min(c(predmean,preds.upper,preds.lower,LEdiffs),na.rm=TRUE)
ymax <- max(c(predmean,preds.upper,preds.lower,LEdiffs),na.rm=TRUE)

plot(as.vector(gensize.conforming), as.vector(LEdiffs),pch=21,cex=0.8,bg="grey80",xlab="Genome Size (pg)",ylab="Leaf Elongation (cm/day)",bty="n",type="n", ylim=c(ymin,ymax))

n.polygons <- length(loggs.seq)-1
for(polygon in 1:n.polygons){
  polygon(
    x=exp(c(loggs.seq[polygon],loggs.seq[polygon],loggs.seq[polygon+1],loggs.seq[polygon+1])), 
    y=c(preds.lower[polygon],preds.upper[polygon],preds.upper[polygon+1],preds.lower[polygon+1]),
    col="gray85",border=NA)
}

lines(exp(loggs.seq),predmean) # zero effect of genome size on leaf elongation
points(as.vector(gensize.conforming), as.vector(LEdiffs),pch=21,cex=0.8,bg="grey80")
```

7-10-16 playing around with plotting realizations onto the fig4 graphs

```{r}
fgrote<-read.csv("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Finalgrowthchamberanalysis/grote_Indexmoms_groundedLL_d16NA.csv")
tmp1<- read.csv("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Grote_stats/Gridcomparison/gamma.chain2.ind4.8mom0_indcellsize.csv")
tmp2<- read.csv("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Grote_stats/Gridcomparison/gamma.chain1.ind4.8mom0_indcellsize.csv")
tmp1 <- mcmc(tmp1,thin=500)
tmp2 <- mcmc(tmp2,thin=500)
#samples2 <- mcmc.list(chain1=tmp1,chain2=tmp2)
samples<-rbind(tmp1,tmp2)

colnames(samples)

#plot(samples.mat[,"tao.GS"],samples.mat[,"LE.coeffs.143."])


#Set up plotting boundaries of the graph
loggs.seq <- seq(from=min(log(fgrote$genomesize)),to=max(log(fgrote$genomesize)), by=0.001)

loggs.mat <- cbind(rep(1,length(loggs.seq)),rep(1,length(loggs.seq)), loggs.seq) #matrix with 3 columns, first 2 columns are =1 to pick up total (sum of individual and maternal or z_ll,i and z_ll,mom(i)) random effect for leaf length and leaf elongation, and the third is a sequence for the axis of log genome size

taoll.positions <- str_detect(colnames(samples),"tao.LL*")
taoll <- samples[,colnames(samples)[taoll.positions]]

lecoef.positions <- str_detect(colnames(samples),"LE.coeffs*")
lecoef <- samples[,colnames(samples)[lecoef.positions]]

#placeholder

all.curves <- 

sample.mean <- ()

predmean <- exp(loggs.mat %*% coef.means) #matrix math for the central tendency curve


leaflength.positions <- str_detect(names(fgrote), "leaf3diff*")#change this to leaf3diff*
leaflength <- fgrote[,names(fgrote)[leaflength.positions]]

#data frames needed for figure generation
gensize.conforming <- matrix(fgrote$genomesize,nrow=nrow(fgrote),ncol=3,byrow=FALSE)
fgrote$leaf3diff10 <- fgrote$leaf3.1 - fgrote$leaf3.0
fgrote$leaf3diff21 <- fgrote$leaf3.2 - fgrote$leaf3.1
fgrote$leaf3diff32 <- fgrote$leaf3.3 - fgrote$leaf3.2
LEdiffs <- as.matrix(cbind(fgrote$leaf3diff10,fgrote$leaf3diff21,fgrote$leaf3diff32))


gs.params <- samples[,c("U.z.indivLL","tao.GS")]
gs.params <- cbind(log(samples$U.z.indivLE),gs.params)#collect the realizations (intercepts) that correspond to the effect being displayed
coef.means <- apply(gs.params, MAR=2, FUN="mean") #take the posterior mean for the central tendency curve

#gs.params needs to have columns for each of the realization terms, might include average(z.indivll+z.momll) or might have other terms.  need to add in a rep(1,length(loggs.seq) for each additional terms

predmean <- exp(loggs.mat %*% coef.means) #matrix math for the central tendency curve

#making of the credibility bands

allpreds <- as.matrix(gs.params) %*% t(loggs.mat) #evaluate the linear predictor for every realization
allpreds <- exp(allpreds) 
#dim(allpreds)
preds.upper <- apply(allpreds,MAR=2,FUN="quantile",probs=c(0.975)) #upper bounds of our predictor
preds.lower <- apply(allpreds,MAR=2,FUN="quantile",probs=c(0.025))

ymin <- min(c(predmean,preds.upper,preds.lower,LEdiffs),na.rm=TRUE)
ymax <- max(c(predmean,preds.upper,preds.lower,LEdiffs),na.rm=TRUE)

plot(as.vector(gensize.conforming), as.vector(LEdiffs),pch=21,cex=0.8,bg="grey80",xlab="Genome Size (pg)",ylab="Leaf Elongation (cm/day)",bty="n",type="n", ylim=c(ymin,ymax))

n.polygons <- length(loggs.seq)-1
for(polygon in 1:n.polygons){
  polygon(
    x=exp(c(loggs.seq[polygon],loggs.seq[polygon],loggs.seq[polygon+1],loggs.seq[polygon+1])), 
    y=c(preds.lower[polygon],preds.upper[polygon],preds.upper[polygon+1],preds.lower[polygon+1]),
    col="gray85",border=NA)
}

lines(exp(loggs.seq),predmean) # zero effect of genome size on leaf elongation
points(as.vector(gensize.conforming), as.vector(LEdiffs),pch=21,cex=0.8,bg="grey80")
```


Genome size by stomatal cell size.  

```{r}
stoma.positions <- str_detect(names(fgrote),"stom*")
stoma <- fgrote[,names(fgrote)[stoma.positions]]
stoma <- 0.0001 * stoma #convert mircons to centimeters

gensize.conforming <- matrix(fgrote$genomesize,nrow=nrow(fgrote),ncol=61,byrow=FALSE) #61 columns for the total number of measures of stoma


loggs.seq <- seq(from=min(log(fgrote$genomesize)),to=max(log(fgrote$genomesize)), by=0.001)

loggs.mat <- cbind(rep(1,length(loggs.seq)), loggs.seq) #DESIGN matrix with 3 columns, first 2 columns are =1 to pick up individual random effect for leaf length and leaf elongation, and the third is a sequence for the axis of log genome size

cs.params <- samples[,c("U.z.indivCS","beta.GS")] #PARAMETER matrix
coef.means <- apply(cs.params, MAR=2, FUN="mean") #take the posterior mean for the central tendency curve

predmean <- exp(loggs.mat %*% coef.means) #matrix math for the central tendency curve

allpreds <- as.matrix(cs.params) %*% t(loggs.mat) #evaluate the linear predictor for every realization
allpreds <- exp(allpreds) 
#dim(allpreds)
preds.upper <- apply(allpreds,MAR=2,FUN="quantile",probs=c(0.975)) #upper bounds of our predictor
preds.lower <- apply(allpreds,MAR=2,FUN="quantile",probs=c(0.025))

ymin <- min(c(predmean,preds.upper,preds.lower,as.matrix(stoma)),na.rm=TRUE)
ymax <- max(c(predmean,preds.upper,preds.lower,as.matrix(stoma)),na.rm=TRUE)


plot(as.vector(gensize.conforming), as.vector(as.matrix(stoma)),pch=".",cex=2,xlab="Genome Size (pg)",ylab="Stomatal Cell Size (cm)",bty="n",type="n")

n.polygons <- length(loggs.seq)-1
for(polygon in 1:n.polygons){
  polygon(
    x=exp(c(loggs.seq[polygon],loggs.seq[polygon],loggs.seq[polygon+1],loggs.seq[polygon+1])), 
    y=c(preds.lower[polygon],preds.upper[polygon],preds.upper[polygon+1],preds.lower[polygon+1]),
    col="gray85",border=NA)
}

lines(exp(loggs.seq),predmean) # zero effect of genome size on leaf elongation
points(as.vector(gensize.conforming), as.vector(as.matrix(stoma)),pch=".",cex=2,bg="grey80")
```

LME4 was abandoned as it could not simultaneously estimate the many components of this model.  Still to do is the inclusion of more error terms to widen the confidence bands.

?dgamma

x <- (0:100 + 0.01) / 50

plot(x,dgamma(x,shape=1, scale=1/1),type="l")
lines(x,dgamma(x,shape=5, scale=1/5),col="red")
mean(rgamma(10000,shape=1, scale=1/1))



