---
title: "Growthchamber_analyses"
output: html_document
---

Load packages

```{r}
library(rjags)
library(stringr)
library(mcmcplots)
```

Rerun with no momGS, nothing should change, we are being more clear that GS is observed, not a thing to predict.  Removed trackers on terms that no longer exist.

```{r}
fgrote<-read.csv("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Grote_stats/grote_Indexmoms_groundedLL_d16NA.csv")

#correct for differences, wont be needed in the next data
#fgrote$leaf3diff10 <- fgrote$leaf3.1 - fgrote$leaf3.0
#fgrote$leaf3diff21 <- fgrote$leaf3.2 - fgrote$leaf3.1
#fgrote$leaf3diff32 <- fgrote$leaf3.3 - fgrote$leaf3.2

#make matrices with data to feed into JAGS
stoma.positions <- str_detect(names(fgrote),"stom*")
stoma <- fgrote[,names(fgrote)[stoma.positions]]
stoma <- 0.0001 * stoma #convert mircons to centimeters
leaflength.positions <- str_detect(names(fgrote), "leaf3diff*")#change this to leaf3diff*
leaflength <- fgrote[,names(fgrote)[leaflength.positions]]
n.cells <- apply(!is.na(stoma), MAR=1, FUN="sum")

DAGLL.jags <- jags.model("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Grote_stats/DAG_twomediator_NomomGS.txt", list(leaflength=log(leaflength),stoma=log(stoma),n.mothers=max(fgrote$momindex),n.indiv=dim(fgrote)[1],n.cells=n.cells,n.times=3,mother.index=fgrote$momindex,GS=log(fgrote$genomesize)),n.chains=2)#n.times=3

update(DAGLL.jags, n.iter=100000)

DAGLL.samples <- coda.samples(DAGLL.jags, c("beta.GS","tao.GS","GS.contrast","U.z.indivCS","U.z.indivLL","U.z.indivLE","sd.motherStCS","sd.motherLL","sd.motherLE","sd.indivCS","sd.indivLL","sd.indivLE","sd.StCS","sd.LL"),n.iter=1000000, thin=1000)

mcmcplot(DAGLL.samples)

chain1<- as.matrix(DAGLL.samples[[1]])
chain2<- as.matrix(DAGLL.samples[[2]])
write.csv(chain1,"~/Desktop/DAGLLedit.jags.1mil.chain1.csv")
write.csv(chain2,"~/Desktop/DAGLLedit.jags.1mil.chain2.csv")
```

Think about the effect of supralinear growth in the explanation of things!

Looking at Chains.

```{r}
tmp1 <- read.csv("~/Desktop/DAGLL.jags.1mil.chain1.csv")
tmp2<- read.csv("~/Desktop/DAGLL.jags.1mil.chain2.csv")

data<- rbind(tmp1,tmp2)

plot(density(data$U.z.indivLE))

```

Graph making.  Goal is to have plots that include the betas.

```{r}
#read in JAGS samples
tmp1<- read.csv("~/Desktop/DAGLLedit.jags.1mil.chain1.csv")
tmp2<- read.csv("~/Desktop/DAGLLedit.jags.1mil.chain2.csv")
samples<- rbind(tmp1,tmp2)

#read in observations
fgrote<-read.csv("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Grote_stats/grote_Indexmoms_groundedLL_d16NA.csv")

#remember variables
stoma.positions <- str_detect(names(fgrote),"stom*")
stoma <- fgrote[,names(fgrote)[stoma.positions]]
stoma <- 0.0001 * stoma #convert mircons to centimeters
leaflength.positions <- str_detect(names(fgrote), "leaf3.*")
leaflength <- fgrote[,names(fgrote)[leaflength.positions]]
n.cells <- apply(!is.na(stoma), MAR=1, FUN="sum")

#correct for differences, wont be needed in the next data
fgrote$leaf3diff10 <- fgrote$leaf3.1 - fgrote$leaf3.0
fgrote$leaf3diff21 <- fgrote$leaf3.2 - fgrote$leaf3.1
fgrote$leaf3diff32 <- fgrote$leaf3.3 - fgrote$leaf3.2
LEdiffs <- as.matrix(cbind(fgrote$leaf3diff10,fgrote$leaf3diff21,fgrote$leaf3diff32))
#emperical derivates, change per day
LEdiffs[which(LEdiffs>12)]

gensize.conforming <- matrix(fgrote$genomesize,nrow=nrow(fgrote),ncol=3,byrow=FALSE)

plot(as.vector(gensize.conforming), as.vector(LEdiffs),pch=21,cex=0.8,bg="grey80",xlab="Genome Size (pg)",ylab="Leaf Elongation (cm/day)",bty="n")

#follow up on the leaf measures that are fractional, dont elongate much

loggs.seq <- seq(from=min(log(fgrote$genomesize)),to=max(log(fgrote$genomesize)), by=0.001)

loggs.mat <- cbind(rep(1,length(loggs.seq)),rep(1,length(loggs.seq)), loggs.seq) #matrix with 3 columns, first 2 columns are =1 to pick up individual random effect for leaf length and leaf elongation, and the third is a sequence for the axis of log genome size

gs.params <- samples[,c("U.z.indivLL","tao.GS")]
gs.params <- cbind(log(samples$U.z.indivLE),gs.params)#collect the realizations (intercepts) that correspond to the effect being displayed
coef.means <- apply(gs.params, MAR=2, FUN="mean") #take the posterior mean for the central tendency curve

#gs.params needs to have columns for each of the realization terms, might include average(z.indivll+z.momll) or might have other terms.  need to add in a rep(1,length(loggs.seq) for each additional terms

predmean <- exp(loggs.mat %*% coef.means) #matrix math for the central tendency curve

#making of the credibility bands

allpreds <- as.matrix(gs.params) %*% t(loggs.mat) #evaluate the linear predictor for every realization
allpreds <- exp(allpreds) 
#dim(allpreds)
preds.upper <- apply(allpreds,MAR=2,FUN="quantile",probs=c(0.975)) #upper bounds of our predictor
preds.lower <- apply(allpreds,MAR=2,FUN="quantile",probs=c(0.025))

ymin <- min(c(predmean,preds.upper,preds.lower,LEdiffs),na.rm=TRUE)
ymax <- max(c(predmean,preds.upper,preds.lower,LEdiffs),na.rm=TRUE)

plot(as.vector(gensize.conforming), as.vector(LEdiffs),pch=21,cex=0.8,bg="grey80",xlab="Genome Size (pg)",ylab="Leaf Elongation (cm/day)",bty="n",type="n", ylim=c(ymin,ymax))

n.polygons <- length(loggs.seq)-1
for(polygon in 1:n.polygons){
  polygon(
    x=exp(c(loggs.seq[polygon],loggs.seq[polygon],loggs.seq[polygon+1],loggs.seq[polygon+1])), 
    y=c(preds.lower[polygon],preds.upper[polygon],preds.upper[polygon+1],preds.lower[polygon+1]),
    col="gray85",border=NA)
}

lines(exp(loggs.seq),predmean) # zero effect of genome size on leaf elongation
points(as.vector(gensize.conforming), as.vector(LEdiffs),pch=21,cex=0.8,bg="grey80")


```

#chunk for running the 6/1 model of JAGS

```{r}
fgrote<-read.csv("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Grote_stats/grote_Indexmoms_groundedLL_d16NA.csv")

#correct for differences, wont be needed in the next data
#fgrote$leaf3diff10 <- fgrote$leaf3.1 - fgrote$leaf3.0
#fgrote$leaf3diff21 <- fgrote$leaf3.2 - fgrote$leaf3.1
#fgrote$leaf3diff32 <- fgrote$leaf3.3 - fgrote$leaf3.2

#make matrices with data to feed into JAGS
stoma.positions <- str_detect(names(fgrote),"stom*")
stoma <- fgrote[,names(fgrote)[stoma.positions]]
stoma <- 0.0001 * stoma #convert mircons to centimeters
leaflength.positions <- str_detect(names(fgrote), "leaf3diff*")#change this to leaf3diff*
leaflength <- fgrote[,names(fgrote)[leaflength.positions]]
n.cells <- apply(!is.na(stoma), MAR=1, FUN="sum")

DAGLL.jags <- jags.model("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Grote_stats/DAG_twomediator_NomomGS.txt", list(leaflength=log(leaflength),stoma=log(stoma),n.mothers=max(fgrote$momindex),n.indiv=dim(fgrote)[1],n.cells=n.cells,n.times=3,mother.index=fgrote$momindex,GS=log(fgrote$genomesize)),n.chains=2)#n.times=3

update(DAGLL.jags, n.iter=20000)

#DAGLL.samples <- coda.samples(DAGLL.jags, c("beta.GS","tao.GS","GS.contrast","U.z.indivCS","U.z.indivLL","U.z.indivLE","sd.motherStCS","sd.motherLL","sd.motherLE","sd.indivCS","sd.indivLL","sd.indivLE","sd.StCS","sd.LL","U.tao.LL","U.LE.coeffs"),n.iter=10000, thin=100)

DAGLL.samples <- coda.samples(DAGLL.jags, c("beta.GS","tao.GS","GS.contrast","U.z.indivCS","U.z.indivLL","U.z.indivLE","U.tao.LL","U.LE.coeffs","U.StCS.coeffs"),n.iter=100000, thin=100)


mcmcplot(DAGLL.samples)

chain1<- as.matrix(DAGLL.samples[[1]])
chain2<- as.matrix(DAGLL.samples[[2]])
write.csv(chain1,"~/Desktop/DAGLL10prior.jags.1mil.chain1.csv")
write.csv(chain2,"~/Desktop/DAGLL10prior.jags.1mil.chain2.csv")
```


#plot(density(samples$sd.motherLE))
#plot(density(samples$sd.motherLL))

6/2 making graphs

```{r}
#read in JAGS samples
tmp1<- read.csv("~/Desktop/DAGLLgraphs.jags.1mil.chain1.csv")
tmp2<- read.csv("~/Desktop/DAGLLgraphs.jags.1mil.chain2.csv")
samples<- rbind(chain1,chain2)

#read in observations
fgrote<-read.csv("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Grote_stats/grote_Indexmoms_groundedLL_d16NA.csv")

#remember variables
stoma.positions <- str_detect(names(fgrote),"stom*")
stoma <- fgrote[,names(fgrote)[stoma.positions]]
stoma <- 0.0001 * stoma #convert mircons to centimeters
leaflength.positions <- str_detect(names(fgrote), "leaf3.*")
leaflength <- fgrote[,names(fgrote)[leaflength.positions]]
n.cells <- apply(!is.na(stoma), MAR=1, FUN="sum")

#correct for differences, wont be needed in the next data
fgrote$leaf3diff10 <- fgrote$leaf3.1 - fgrote$leaf3.0
fgrote$leaf3diff21 <- fgrote$leaf3.2 - fgrote$leaf3.1
fgrote$leaf3diff32 <- fgrote$leaf3.3 - fgrote$leaf3.2
LEdiffs <- as.matrix(cbind(fgrote$leaf3diff10,fgrote$leaf3diff21,fgrote$leaf3diff32))
#emperical derivates, change per day
LEdiffs[which(LEdiffs>12)]

gensize.conforming <- matrix(fgrote$genomesize,nrow=nrow(fgrote),ncol=3,byrow=FALSE)

plot(as.vector(gensize.conforming), as.vector(LEdiffs),pch=21,cex=0.8,bg="grey80",xlab="Genome Size (pg)",ylab="Leaf Elongation (cm/day)",bty="n")

#follow up on the leaf measures that are fractional, dont elongate much

loggs.seq <- seq(from=min(log(fgrote$genomesize)),to=max(log(fgrote$genomesize)), by=0.001)

loggs.mat <- cbind(rep(1,length(loggs.seq)),rep(1,length(loggs.seq)), loggs.seq) #DESIGN matrix with 3 columns, first 2 columns are =1 to pick up individual random effect for leaf length and leaf elongation, and the third is a sequence for the axis of log genome size

gs.params <- samples[,c("U.tao.LL","tao.GS")] #PARAMETER matrix
gs.params <- cbind(log(samples[,"U.LE.coeffs"]),gs.params)#collect the realizations (intercepts) that correspond to the effect being displayed
colnames(gs.params)[1] <- "U.LE.coeffs"
coef.means <- apply(gs.params, MAR=2, FUN="mean") #take the posterior mean for the central tendency curve

#gs.params needs to have columns for each of the realization terms, might include average(z.indivll+z.momll) or might have other terms.  need to add in a rep(1,length(loggs.seq) for each additional terms

predmean <- exp(loggs.mat %*% coef.means) #matrix math for the central tendency curve

#making of the credibility bands

allpreds <- as.matrix(gs.params) %*% t(loggs.mat) #evaluate the linear predictor for every realization
allpreds <- exp(allpreds) 
#dim(allpreds)
preds.upper <- apply(allpreds,MAR=2,FUN="quantile",probs=c(0.975)) #upper bounds of our predictor
preds.lower <- apply(allpreds,MAR=2,FUN="quantile",probs=c(0.025))

ymin <- min(c(predmean,preds.upper,preds.lower,LEdiffs),na.rm=TRUE)
ymax <- max(c(predmean,preds.upper,preds.lower,LEdiffs),na.rm=TRUE)

plot(as.vector(gensize.conforming), as.vector(LEdiffs),pch=21,cex=0.8,bg="grey80",xlab="Genome Size (pg)",ylab="Leaf Elongation (cm/day)",bty="n",type="n", ylim=c(ymin,ymax))

n.polygons <- length(loggs.seq)-1
for(polygon in 1:n.polygons){
  polygon(
    x=exp(c(loggs.seq[polygon],loggs.seq[polygon],loggs.seq[polygon+1],loggs.seq[polygon+1])), 
    y=c(preds.lower[polygon],preds.upper[polygon],preds.upper[polygon+1],preds.lower[polygon+1]),
    col="gray85",border=NA)
}

lines(exp(loggs.seq),predmean) # zero effect of genome size on leaf elongation
points(as.vector(gensize.conforming), as.vector(LEdiffs),pch=21,cex=0.8,bg="grey80")

```

#FINDME

```{r}
samples<- rbind(chain1,chain2)
fgrote<-read.csv("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Grote_stats/grote_Indexmoms_groundedLL_d16NA.csv")

stoma.positions <- str_detect(names(fgrote),"stom*")
stoma <- fgrote[,names(fgrote)[stoma.positions]]
stoma <- 0.0001 * stoma #convert mircons to centimeters

gensize.conforming <- matrix(fgrote$genomesize,nrow=nrow(fgrote),ncol=61,byrow=FALSE) #61 columns for the total number of measures of stoma

loggs.seq <- seq(from=min(log(fgrote$genomesize)),to=max(log(fgrote$genomesize)), by=0.001)

loggs.mat <- cbind(rep(1,length(loggs.seq)), loggs.seq) #DESIGN matrix with 3 columns, first 2 columns are =1 to pick up individual random effect for leaf length and leaf elongation, and the third is a sequence for the axis of log genome size

cs.params <- samples[,c("U.z.indivCS","beta.GS")] #PARAMETER matrix
coef.means <- apply(cs.params, MAR=2, FUN="mean") #take the posterior mean for the central tendency curve

predmean <- exp(loggs.mat %*% coef.means) #matrix math for the central tendency curve

allpreds <- as.matrix(cs.params) %*% t(loggs.mat) #evaluate the linear predictor for every realization
allpreds <- exp(allpreds) 
#dim(allpreds)
preds.upper <- apply(allpreds,MAR=2,FUN="quantile",probs=c(0.975)) #upper bounds of our predictor
preds.lower <- apply(allpreds,MAR=2,FUN="quantile",probs=c(0.025))

ymin <- min(c(predmean,preds.upper,preds.lower,as.matrix(stoma)),na.rm=TRUE)
ymax <- max(c(predmean,preds.upper,preds.lower,as.matrix(stoma)),na.rm=TRUE)


plot(as.vector(gensize.conforming), as.vector(as.matrix(stoma)),pch=".",cex=2,xlab="Genome Size (pg)",ylab="Stomatal Cell Size (cm)",bty="n",type="n")

n.polygons <- length(loggs.seq)-1
for(polygon in 1:n.polygons){
  polygon(
    x=exp(c(loggs.seq[polygon],loggs.seq[polygon],loggs.seq[polygon+1],loggs.seq[polygon+1])), 
    y=c(preds.lower[polygon],preds.upper[polygon],preds.upper[polygon+1],preds.lower[polygon+1]),
    col="gray85",border=NA)
}

lines(exp(loggs.seq),predmean) # zero effect of genome size on leaf elongation
points(as.vector(gensize.conforming), as.vector(as.matrix(stoma)),pch=".",cex=2,bg="grey80")

```

The lme running for cell size.  This can only operate on one variable at a time, so to judge our lme model fits, we have to do cell size and leaf length separately.  The lme returns similar estimates of our individual and mom random effects, and even the residual of cell size.  However, it shows the identification problem in the beta.gs when we look at the covar between genomesize and intercept.  The covar is -1, showing that our estimate of beta.gs in lme, which is markedly different from our JAGS model, is not to be believed because it has an identification problem.

```{r}
#install.packages("lme4")
library(lme4)
library(nlme)
library(stringr)

fgrote<-read.csv("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Grote_stats/grote_Indexmoms.csv")

stoma.positions <- str_detect(names(fgrote),"stom*")
stoma <- fgrote[,names(fgrote)[stoma.positions]]
stoma <- 0.0001 * stoma #convert mircons to centimeters


cslme4 <- fgrote[,c("fullid","momindex","genomesize")]
cslme4 <- data.frame(cslme4,stoma)

cslme4.long <- reshape(cslme4, varying= list(4:64), idvar="fullid",direction="long")
names(cslme4.long) <- c("fullid","momindex","genomesize","replicate","cellsize")
sum(complete.cases(cslme4.long))

cslme4.long <- subset(cslme4.long, subset = !is.na(cslme4.long$cellsize))

#frequentist model for log cell size with a standalone intercept; intercept exists permanently, unless we specify otherwise, which is our ultimate plan; has coefficient of log GS and random effects of individuals within moms
m1.cs <- lme(log(cellsize) ~ log(genomesize), random= ~ 1|momindex/fullid, data=cslme4.long )

#second model that removes the standalone intercept; -1 is the syntax to get rid of the intercept term
m2.cs <- lme(log(cellsize) ~ -1 + log(genomesize), random= ~ 1|momindex/fullid, data=cslme4.long )
summary(m2.cs)
#residual for cell size standard deviation .167=.167, roughly equal between bayes/freq
#individual random effect standard dev .125 ~ 0.095 (freq ~ bayes)
#mother cell size random effect stand dev 0.077 ~ 0.05 (freq ~ bayes)

ranef(m2.cs, level=2)
#THESE ARE BLUPS.  they are predictions of the individual level effects, and we want to see if they behave similarly to the estimates of individual effects from bayesians.
#we are talking about where the info in the intercepts go.

#mom random effects, centered at 0
blups <- as.data.frame(ranef(m2.cs, level=1))
colnames(blups)[1] <- "Intercept"
tmp<- as.numeric(blups$Intercept)
plot(density(tmp))

#individual random effects, centered at 0
blups <- as.data.frame(ranef(m2.cs, level=2))
colnames(blups)[1] <- "Intercept"
tmp<- as.numeric(blups$Intercept)
plot(density(tmp))
#with both of these centered at 0, we can conclude that it is unlikely that our frequentist and bayesian models can be compared directly.  so can't build confidence from independent verification, but we gave it a go



chain1<-read.csv("~/Desktop/DAGLLedit.jags.1mil.chain1.csv")
chain2<-read.csv("~/Desktop/DAGLLedit.jags.1mil.chain2.csv")
samples<- rbind(chain1,chain2)
summary(samples)
mcmcplot(samples)
cor(vcov(m1.cs))
```

```{r}
leaflength.positions <- str_detect(names(fgrote), "leaf3.*")#change this to leaf3diff*
leaflength <- fgrote[,names(fgrote)[leaflength.positions]]

fgrote$ll.1 <- fgrote$leaf3.1 - fgrote$leaf3.0
fgrote$ll.2 <- fgrote$leaf3.2 - fgrote$leaf3.0
fgrote$ll.3 <- fgrote$leaf3.3 - fgrote$leaf3.0
leaflengths <- cbind(fgrote$ll.1, fgrote$ll.2, fgrote$ll.3)


ll.lme <- fgrote[,c("fullid","momindex","genomesize")]
ll.lme <- data.frame(ll.lme,leaflengths)
ll.lme.long <- reshape(ll.lme, varying= list(4:6), idvar="fullid",direction="long")


names(ll.lme.long) <- c("fullid","momindex","genomesize","time","length")
sum(complete.cases(ll.lme.long))

ll.lme.long <- subset(ll.lme.long, subset = !is.na(ll.lme.long$length))

lmeControl(maxIter = 500, msVerbose = TRUE)
m.ll <- lme(log(length) ~ log(genomesize) + log(time), random= ~ 1 + log(time) | momindex/fullid, data=ll.lme.long)


```



####END OF DOCUMENT

Chunk will read in the stats, prepare the JAGS model, and run the model.  Number of iterations should be changed.

```{r}
fgrote<-read.csv("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Grote_stats/grote_Indexmoms.csv")

#make matrices with data to feed into JAGS
stoma.positions <- str_detect(names(fgrote),"stom*")
stoma <- fgrote[,names(fgrote)[stoma.positions]]
stoma <- 0.0001 * stoma #convert mircons to centimeters
leaflength.positions <- str_detect(names(fgrote), "leaf3.*")#change this to leaf3diff*
leaflength <- fgrote[,names(fgrote)[leaflength.positions]]
n.cells <- apply(!is.na(stoma), MAR=1, FUN="sum")

DAGLL.jags <- jags.model("~/Documents/Projects/Genome_Size_Analysis/Github_ParallelGS/Mexicana/Grote_stats/DAG_twomediator_forLL.txt", list(leaflength=log(leaflength),stoma=log(stoma),n.mothers=max(fgrote$momindex),n.indiv=dim(fgrote)[1],n.cells=n.cells,n.times=4,mother.index=fgrote$momindex,GS=log(fgrote$genomesize)),n.chains=2)#n.times=3

update(DAGLL.jags, n.iter=1000)

DAGLL.samples <- coda.samples(DAGLL.jags, c("beta.GS","gamma.GS","GS.contrast","U.z.motherGS","U.z.indivCS","U.z.indivLL","U.z.indivLE","sd.motherStCS","sd.motherGS","sd.motherLL","sd.motherLE","sd.GS","sd.indivCS","sd.indivLL","sd.indivLE","sd.StCS","sd.LL"),n.iter=50000, thin=200)
#arra
```
